---
title: "Chapter 1: shiny"
author: "Jumping Rivers"
date: "`r Sys.Date()`"
output: ioslides_presentation
runtime: shiny
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 4 Slides}
-->

# Setup
## Setup

Please do the following:

Basics:

```{r eval=FALSE}
install.packages("drat")
drat::addRepo("jr-packages")
install.packages("jrShiny")

# start a new Rstudio project (optional)
jrShiny::create_project()
```

You should see two new directories `exercises` and `original`. We will work on `exercises` leaving original as a reference.

Advanced:

Download or clone the repostiory - 

```{r echo=FALSE, message=FALSE}
library(shiny)
data(movies, package="jrShiny")
```

# Shiny
## Shiny

* [Shiny](https://goo.gl/tzZs3T) is a R package that provides a web 
application framework for R
* Fast and simple development for interactive graphics.
* See also
* http://shiny.rstudio.com/gallery
* http://www.showmeshiny.com/
* http://shiny.rstudio.com/tutorial/
* http://shiny.rstudio.com/articles/
* Every shiny app is maintained by a computer running R

## Shiny

* Provides functions which allow automatic bindings between input and output
* Build interactive web applications using R
* Gives an interface between R programming and JavaScript/html output
* No knowledge of JavaScript/html required

## First app: `Rmd`

```
---
title: "First app"
runtime: shiny
---
```

# Input: control widgets
## Input: control widgets

```{r echo=FALSE}
selectInput("movie_type1", # unique id 
            label = "Movie genre", # Text for web
            c("Romance", "Action", "Animation"))
checkboxGroupInput("id1", label = "Checkbox", c("Romance", "Action", "Animation"))
```
Other [Widgets](http://shiny.rstudio.com/gallery/widget-gallery.html) Gallery

## Widgets
This
```{r results="hide"}
selectInput("movie_type", # unique id 
            label = "Movie genre", # Text for web
            c("Romance", "Action", "Animation"))
```
gives
```
<div class="form-group shiny-input-container">
<label class="control-label" for="movie_type">Movie genre</label>
<div>
<select id="movie_type">
<option value="Romance" selected>Romance</option>
<option value="Action">Action</option>
<option value="Animation">Animation</option></select>
<script type="application/json" data-for="movie_type" 
data-nonempty="">{}</script>
</div>
</div>
```
# Exercise 2
## Exercise 2

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```

# Rendered outputs
## Rendered outputs

* Use the values in the shiny widgets to dynamically change
plots and tables.
* When we create a shiny widget, the value of the widget is
bound to `input$inputId`
* This object is a _reactive_ value

> What code will the server run? Code that builds something out of `reactive` values

> When will the code run? When the _reactive_ value changes.


# Exercise 3
## Exercise 3 

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```
## Rendered output

* Rendered (outputs) respond whenever a `reactive` value changes. 
* For example, the scatter plot when we were exploring the occurrence of names.

Function | Output type
---------|------------
`renderPlot` | R graphics output
`renderPrint` | printed output
`renderTable`	| Data frame, matrix
`renderText` | Character vectors

## Example: movies

```{r eval=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})
```
Or
```{r eval=FALSE}
renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

# Exercise 4
## Exercise 4

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```

# Reactive programming
## Reactive programming

* __render__ functions == reactive programming
* Standard R session

```{r}
x = 1
print(x)
x = 2 ## The print statement above is not updated.
```

* Shiny gives the illustration that `print(x)` is updated.

## Reactive programming

*   A naive way of doing this is to constantly check if any objects have changed.
* Doesn't scale 
* shiny just checks key objects. If these objects have changed, the necessary changes
are propagated through.
* When we include `input$X` within a `render`
- indicates that we should rerun render when `input$X` changes

> A reactive expression is an R expression that uses widget input and returns a
value.

## Problem

This is inefficient

```{r eval=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})
renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

## Solution: `reactive`

```{r eval=FALSE}
sub_movies  = reactive(movies[movies[input$movie_type]==1,])
```
Or 
```{r eval=FALSE}
sub_movies = reactive({
  type = movies[,input$movie_type] == 1
  movies[type,]
})
```
Gives

```{r eval=FALSE}
renderText(nrow(sub_movies()))
renderPlot(hist(sub_movies()$length))
```

# Exercise 5
## Exercise 5

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```

# `eventReactive`
## Problem

```{r eval=FALSE}
selectInput("movie_genre", label = "Movie genre", 
            c("Romance", "Action", "Animation"))
actionButton("plot", "Plot it now!!")

sub_movies  = reactive(movies[movies[input$movie_genre]==1,])
#??
```

## Solution: `actionButton` and `actionLink`

```{r echo=TRUE}
selectInput("movie_genre", label = "Movie genre", 
            c("Romance", "Action", "Animation"))
actionButton("plot", "Plot it now!!")

sub_movies1 = eventReactive(input$plot,{
  type = movies[,input$movie_genre] == 1
  movies[type,]
})
renderPlot(hist(sub_movies1()$length))
```

# Exercise 6
## Exercise 6

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```

# The `observeEvent` function
## Problem

```{r}
actionButton("romance", "Romance")
actionButton("action", "Action")
```

* The `eventReactive` approach doesn't work. 

## Solution: `reactiveValues` and `observeEvent`

```{r eval=FALSE}
rvs = reactiveValues(data=movies)
```

Next we monitor the buttons for changes, and update `rvs` as needed

```{r eval=FALSE}
observeEvent(input$romance, {
  rvs$data = {
    type = movies[,input$movie_type] == 1
    movies[type,]
  }
})
observeEvent(input$action, {
  rvs$data = {
    type = movies[,input$movie_type] == 1
    movies[type,]
  }
})
renderPlot(hist(rvs$data[, "length"]))
```

# Exercise 7
## Exercise 7

```{r message=FALSE, eval=FALSE}
vignette("chapter1", package="jrShiny")
```

# The `isolate` function
## Problem

```{r eval=FALSE}
sliderInput("n", "Sample size", 10, 500, 100)
actionButton("romance", "Romance")
actionButton("action", "Action")
```


```{r eval=FALSE}
observeEvent(input$romance, {
  m =  movies[movies[input$movie_type]==1,]
  rows = sample(1:nrow(m), input$n)
  rvs$data = movies[rows,]
})
# Similar for Action
renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", input$n)))
```

## Problem

```{r shiny1, echo=FALSE}
sliderInput("n_iso", "Sample size", 10, 500, 100)
actionButton("romance_iso", "Romance")
actionButton("action_iso", "Action")
rvs = reactiveValues(data=movies)
observeEvent(input$romance_iso, {
  m =  movies[movies[,"Romance"]==1,]
  rows = sample(1:nrow(m), input$n_iso)
  rvs$data = m[rows,]
})

observeEvent(input$action_iso, {
  m =  movies[movies[,"Action"]==1,]
  rows = sample(1:nrow(m), input$n_iso)
  rvs$data = m[rows,]
})
renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", input$n_iso)))
```

## Solution: `isolate`

```{r eval=FALSE}
renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", isolate(input$n))))
```

## Solution: `isolate`

```{r echo=FALSE}
sliderInput("n_iso1", "Sample size", 10, 500, 100)
actionButton("romance_iso1", "Romance")
actionButton("action_iso1", "Action")
rvs = reactiveValues(data=movies)
observeEvent(input$romance_iso1, {
  m =  movies[movies[,"Romance"]==1,]
  rows = sample(1:nrow(m), input$n_iso1)
  rvs$data = m[rows,]
})

observeEvent(input$action_iso1, {
  m =  movies[movies[,"Action"]==1,]
  rows = sample(1:nrow(m), input$n_iso1)
  rvs$data = m[rows,]
})

renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", isolate(input$n_iso1))))
```

## Flexdashboard

* See if you can recreate [this](https://bookdown.org/csgillespie/shiny_components_reactive/)
* The basic structure has been started for you in `chapter1/flexstarter.Rmd`